%*******************************************************
% Acknowledgements
%*******************************************************
\pdfbookmark[1]{Acknowledgements}{acknowledgements}


\bigskip


\chapter*{Acknowledgements}

A great many people supported me during this journey. This work would not have been successfully completed without the help, support and trust of mentors, friends and family. Foremost, I would like to acknowledge the guidance from my supervisors, Associate Professors Loulin Huang and Russel Pears. 

This research was financially supported by the Ministry of Business, Innovation and Employment (MBIE) under contract number: AITX0901. I was very  fortunate to receive wonderful support from the staff at MBIE who helped me to transition from Massey University to Auckland University of Technology (AUT). 

Nga mihi nui m{\={o} \={o} man\={a}kitanga, ten\={a} r\={a}w\={a} atu koutou, in appreciation of your kindness many thanks, to all the staff at AUT who have guided me along the way. My sincere thanks to Kath Ravenhall from the School of Engineering, Emma Kelly and the staff from the Research and Innovation Office, Tui O'Sullivan and all those who helped to organise and produce the exhibition on native bees. 

I am indebted to the developers of the medical imaging package FIJI and would especially like to thank Ignacio Arganda--Carreras, who was ever generous with his time and resources.
\bigskip


\begin{flushright}{\slshape    
    ...N\={o} reira, t\={e}n\={a} koutou, t\={e}n\={a} koutou, \={a} t\={e}n\={a} t\={a}tou katoa.} 
\end{flushright}




